\Large \bf FlexAlloc: Dynamic Memory Partitioning\\ between the Corevisor and Hostvisor in SeKVM

### 2023-12-23 paper.tex
Some formatting fixes

We implement this fragmentation by manipulating the iterators used to traverse
the guest VM's stage-2 page table. In particular, the VM's stage-2 page table
is a four-level page table, comprising a page global directory (PGD), a
page upper directory (PUD), a page middle directory (PMD), and multiple page table
entries (PTEs). TODO (Matthew - Page table fragmentation)

@@

TODO (Matthew - DPM)

>>>>

We implement this fragmentation by manipulating the iterators that provide the next useable
page at a given level. These iterators are used when traversing the stage-2 page table in order to record
an additional entry. In the original SeKVM implementation, the pages in each level of the stage-2 page table are
kept contiguous to one another, and never intermixed with pages from other levels.
Each iterator starts at a particular offset into the VM's statically allocated stage-2
page table region, and increments each time a new entry is consumed at that level. The design of these iterators means that they are limited to
only operating within a large, contiguous region. In particular, the PGD iterator operates over a \textit{16 * PAGE\_SIZE} region, the PUD
iterator operates over a \textit{SZ\_2M - (16 * PAGE\_SIZE)} region, and the PMD iterator operates over a \textit{3 * SZ\_2M} region.

In contrast, our design eliminates the need for all of these subregions to be contiguous to
one another and enables fragmentation within each subregion itself. Each iterator can be thought of as a page allocator specific to VM stage-2
page tables, and operates over a pool of eight, non-contiguous 1MB regions. The FlexAlloc iterators
first exhaust all pages from a given 1MB region. Once the current 1MB region is exhausted, FlexAlloc
will jump the iterators to the next 1MB region in the pool. Subsequent traversals of the
stage-2 page tables with the alloc flag enabled will pull pages from the seperate, second
non-contiguous 1M region. By removing the requirement for contiguous 8MB regions, FlexAlloc
enables reliable dynamic allocation of stage-2 page table regions.

We implement two versions of the FlexAlloc iterators. In the first version, we use a single,
generalized iterator shared across all levels of the stage-2 page table. In the SeKVM implementation,
the iterators are structured such that a PUD page will never be sandwiches between two PMD pages.
The first version of the FlexAlloc iterators ignores this semantic and instead is non-discriminatory
in regards to the particular level a page will be used in. However, we eventually decided to stop development of this
version before it was fully functional, in favor of version two which more closely resembles recent changes made to SeKVM's
stage-2 page table allocation scheme.

In the second implementation, we opt to enforce the semantic that pages at each specific level are
allocated from seperate pools, and thus never intermixed with one another. To accomplish this while still maintaining the same pool of non-contiguous 1MB regions,
we split the single, general iterator into nine distinct iterators. One iterator is used
for PUD pages, two iterators are used for PMD pages, and six iterators are used for PTE pages.
The number of iterators assigned to each level is based on the potential maximum number of
pages that could be consumed at that level, along with how many distinct 1MB regions each level
can be allocated from. When a page is allocated at a particular level, the correct iterator is chosen based on
its assigned level and whether or not its preceding iterators have been exhausted.

The naming semantics of the iterators at each level can be confusing, so we feel it is worth
clarifying here. The PGD iterator allocates pages to use as entries inside of a PGD page. This
implies that the PGD iterator, as an example, actually allocates pages that are used in the
subsequent level, which in this case is the PUD. All iterators follow the pattern of allocating
pages for a given level to point to. These are actually pages in the next level
of the stage-2 page tables.

@@

To enable dynamic allocation of EL2 memory in general beyond just the scope of
a VM's stage-2 page tables, we also designed but did not implement
a dynamic page manager (DPM). DPM is a system that exposes an interface for other ``services'' within EL2 to
utilize dynamic allocation for their memory regions, while hiding the implementation of actually receiving memory from
and returning memory to the host. This would be another interesting area of future work.
